# All Data Summary
This file combines all the data across all datasets to calculate overall summary statistics

## Load Libraries
```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

```{r}
# Define column lists for each pollutant group
id_cols <- c("date", "participant_id", "environment", "case")
gas_phase_cols <- c(id_cols, "no", "co", "no2")
particulate_matter_cols <- c(id_cols, "pm1", "pm25", "pm10", "pm1num")
black_carbon_cols <- c(id_cols, "uv_bc1")
```


## Read data
```{r}
path_to_data <- "data_cleaned/"

load(paste0(path_to_data, "cleaned_Olin_modpm.RData"))
modpm_olin <- modpm_df

load(paste0(path_to_data, "cleaned_Olin_modulair.RData"))
modulair_olin <- modulair_df

load(paste0(path_to_data, "cleaned_Olin_cpc.RData"))
cpc_olin <- cpc_df

load(paste0(path_to_data, "cleaned_Olin_aethalometer.RData"))
aethalometer_olin <- aethalometer_df

load(paste0(path_to_data, "cleaned_SCOPE_modulair.RData"))
modulair_scope <- modulair_df

load(paste0(path_to_data, "cleaned_SCOPE_cpc.RData"))
cpc_scope <- cpc_df

load(paste0(path_to_data, "cleaned_Tufts_cpc.RData"))
cpc_tufts <- cpc_df
```

## Combine Data
```{r}
# Combine all Modulair data
modulair_all <- rbind(modulair_olin, modulair_scope)
# Select columns for gas phase pollutants
gas_phase <- modulair_all %>% select(all_of(gas_phase_cols))

# Combine all particulate matter data
particulate_matter <- modulair_all %>% 
  select(all_of(particulate_matter_cols)) %>%
  rbind(modpm_olin)

# Combine Brechtel CPC data, add CPC type to variable
ufp_brechtel <- rbind(cpc_olin, cpc_scope) %>% 
  rename(concent_brechtel = concent)

# Add CPC type to Tufts CPC variable name
ufp_tsi <- cpc_tufts %>% rename(concent_tsi = concent)

# Select columns for black carbon
black_carbon <- aethalometer_olin %>% select(all_of(black_carbon_cols))
```

## Helper Functions: calculate summary statistics
Provided a given grouped dataframe, calculate all relevant summary statistics
```{r}
my_summarise <- function(main_df) {
  main_df %>%
    # Gather dataframe to create single variable for measurements
    pivot_longer(!c(date, environment, participant_id, case), 
                 names_to = "p_type", values_to = "reading") %>%
    # Drop NaN, inf values for summary calculation
    filter(is.finite(reading)) %>%
    group_by(case, environment, p_type) %>%
    summarise(mean = mean(reading),
            median = median(reading), 
            q5 = quantile(reading, probs = 0.05), 
            q25 = quantile(reading, probs = 0.25),
            q75 = quantile(reading, probs = 0.75),
            q95 = quantile(reading, probs = 0.95),
            sd = sd(reading),
            .groups = 'drop')
}
```

## Main Code Run
### Calculate summary statistics acoss all data
```{r}
all_summary <- my_summarise(particulate_matter) %>%
  rbind(my_summarise(gas_phase)) %>%
  rbind(my_summarise(ufp_brechtel)) %>%
  rbind(my_summarise(ufp_tsi)) %>%
  rbind(my_summarise(black_carbon))
```

### Calculate percentage reduction in concentration from no HEPA to HEPA
```{r}
# Spread: Create separate sham, hepa summary statistics columns
summary_wide <- all_summary %>%
  pivot_wider(names_from = case, values_from = mean:sd)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("off")) - 
  select(summary_wide, ends_with("on"))) / 
  select(summary_wide, ends_with("off"))

# Remove (now incorrect) "off" suffix for percentage reduction
colnames(summary_redu) <- sub("_off", "", colnames(summary_redu))

# Add environment, p_type, to redu dataframe 
summary_redu <- summary_wide %>%
  select(environment, p_type) %>%
  cbind(summary_redu) %>%
  drop_na()
```

-- End of summarizing script --

```{r}
## Global variables for plotting function

# Sets order for displaying particles in heatmap
particle_order <- c("PM 1 Conc.", "PM 2.5 Conc.", "PM 10 Conc.", 
                   "PM 1 Count", "UFP (Brechtel)", "UFP (TSI)",
                   "CO Conc.", "NO Conc.", "NO2 Conc.", "BC (UV)")

relevant_pollutants <- c("pm25", "pm1num", "concent_brechtel", "no", "uv_bc1")

stat_name <- c("mean" = "Mean", "median" = "Median", "q5" = "5th Percentile",
               "q25" = "25th Percentile", "q75" = "75th Percentile",
               "q95" = "95th Percentile", "sd" = "Standard Deviation")

plot_heatmap <- function(summary_redu, sum_stat) {
  # Pre-process data before ploting
  summary_redu %>%
    filter(p_type %in% relevant_pollutants) %>%
    
    # Recode for accurate labels
    mutate(p_type = recode(p_type, 
                           "pm1" = "PM 1 Conc.", "pm25" = "PM 2.5 Conc.",
                           "pm10" = "PM 10 Conc.", "pm1num" = "PM 1 Count",
                           "concent_brechtel" = "UFP (Brechtel)",
                           "concent_tsi" = "UFP (TSI)", "co" = "CO Conc.", 
                           "no" = "NO Conc.", "no2" = "NO2 Conc.", 
                           "uv_bc1" = "BC (UV)"),
           environment = recode(environment,
                                "indoor" = "Indoor", 
                                "outdoor" = "Outdoor",
                                "ratio" = "I/O Ratio")) %>%
    
    # Arrange axes in specified order
    mutate(p_type = factor(p_type, levels = particle_order)) %>%

    # Plot heatmap
    ggplot(aes(x = p_type, y = environment)) +
      # Fill with value of stat
      geom_tile(aes(fill = .data[[sum_stat]])) +
      # Print in value of stat
      geom_text(aes(label = paste0(round(.data[[sum_stat]], 1), "%"))) +
    
      # Set color scale
      scale_fill_gradient2(low = "#ff7f7f", high = "#DeF7E9", mid = "white") +
      # Set axis labels, title
      labs(title = paste("Percentage Reductions in", stat_name[sum_stat]), 
           x = "Particle Type", y = "Summary Statistic", 
           fill = "Reduction (%)") +
      # Set theme
      theme_minimal()
}
```

## Plot data as heatmap
```{r}
summary_redu %>% 
  filter(environment != "outdoor") %>%
  plot_heatmap("mean")
```
```{r}
summary_redu %>% 
  filter(environment != "outdoor") %>%
  plot_heatmap("q5")
```

