# Modulair Data Summary
This file is to summarize Modulair Data for Olin deployments

## Load Libraries
```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read data
```{r}
path_to_data <- "data_cleaned/"

load(paste0(path_to_data, "cleaned_Olin_modulair.RData"))
```

## Helper Function: Calculate summary statistics
Provided a given grouped dataframe, calculate all relevant summary statistics
```{r}
my_summarise <- function(main_df, groups) {
  main_df %>%
    group_by(across(all_of(groups))) %>%
    summarise(mean = mean(reading),
            median = median(reading), 
            q5 = quantile(reading, probs = 0.05), 
            q25 = quantile(reading, probs = 0.25),
            q75 = quantile(reading, probs = 0.75),
            q95 = quantile(reading, probs = 0.95),
            sd = sd(reading),
            .groups = 'drop')
}
```

## Main Code Run
### Gather, group, and summarize
```{r}
modulair_df_long <- modulair_df %>%
  # Gather dataframe to create single variable for measurements
  pivot_longer(!c(date, environment, participant_id, case), 
               names_to = "p_type", values_to = "reading") %>%
  # Drop NaN, inf values for summary calculation
  filter(is.finite(reading)) 

# Calculate summary statistics for each participant
participant_summary <- modulair_df_long %>%
  my_summarise(c("participant_id", "case", "environment", "p_type"))
  
# Calculate summary statistics over all the data
overall_summary <- modulair_df_long %>%
  my_summarise(c("case", "environment", "p_type")) %>%
  # Add participant ID column as 'all'
  mutate(participant_id = "all")

# Bind together participant and overall summary statistics
summary <- rbind(participant_summary, overall_summary)
```

### Calculate percentage reduction in concentration from no HEPA to HEPA
```{r}
# Spread: Create separate sham, hepa summary statistics columns
summary_wide <- pivot_wider(summary, names_from = case, values_from = mean:sd)

# Calculate percentage reduction in all summary statistics
summary_redu <- 100*(select(summary_wide, ends_with("off")) - 
  select(summary_wide, ends_with("on"))) / 
  select(summary_wide, ends_with("off"))

# Remove (now incorrect) "off" suffix for percentage reduction
colnames(summary_redu) <- sub("_off", "", colnames(summary_redu))

# Add participant_id, environment, p_type, to redu dataframe 
summary_redu <- summary_wide %>%
  select(participant_id, environment, p_type) %>%
  cbind(summary_redu)
```

```{r}
write_csv(summary, "data_summarized/summary_Olin_modulair_raw.csv")
write_csv(summary_redu, "data_summarized/summary_Olin_modulair_redu.csv")
```

