# Modulair Data Cleaning
This file is to clean the Modulair Data for 4 Olin deployments

## Load Libraries
```{r}
library(tidyverse)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read Data
```{r}
path_to_data <- "data_raw/Olin/modulair/"

# All columns
col_list <- c("timestamp", "pm1", "pm25", "pm10", "no", "co", "no2", "bin0", "bin1", "bin2")

# List of participants with Modulairs
modulair_participants <- c(57301, 57331, 57311, 57341)
```

## Read data
All modulair data was collected by two Modulairs (one indoors, another outdoors) being shuffled between different deployments
```{r}
# Read indoor data
indoor <- read_csv(paste0(path_to_data, "OH_R_indoor.csv"), show_col_types = FALSE) %>%
  select(col_list) %>% mutate(environment = "indoor")

# Read outdoor data
outdoor <- read_csv(paste0(path_to_data, "OH_R_outdoor.csv"), show_col_types = FALSE) %>%
  select(all_of(col_list)) %>% mutate(environment = "outdoor")

# Combine indoor and outdoor data
all_df <- rbind(indoor, outdoor)
```

## Participant, case filtering
Dividing the continuous sensor data between different participants and assigning off (sham) or on (hepa) from field notes
```{r}
# Define lookup table to map "sham" and "hepa" to "off" and "on" respectively
case_lookup <- c(Sham = "off", Hepa = "on")
# Define table to invert the case for the second period
case_inverse <- c(off = "on", on = "off")


# Read field notes, skipping top two rows
time_map <- read_excel("data_raw/Olin/Olin_field_notes.xlsx", range = "A3:I25") %>%
  # Rename and select relevant columns
  rename(participant_id = 1, case_1 = "Filter",
         dt_start_1 = "Air Mon 1 Visit 1", dt_end_1 = "Air Mon 1 Visit 2",
         dt_start_2 = "Air Mon 2 Visit 3", dt_end_2 = "Air Mon 2 Visit 4") %>%
  select(participant_id, case_1, starts_with("dt_")) %>%
  # Convert timezone to Eastern Time
  mutate(across(c(dt_start_1, dt_end_1, dt_start_2, dt_end_2), 
                ~force_tz(.x, tzone = "America/New_York")), 
         # Participant ID to numeric
         participant_id = as.numeric(participant_id),
         # Standardize case to on, off, invert the case in the second period
         case_1 = case_lookup[case_1], case_2 = case_inverse[case_1]) %>%
  # Filter for Modulair participants
  filter(participant_id %in% modulair_participants) %>%
  # Stretch to create case dataframe
  pivot_longer(cols = -participant_id, names_to = c(".value"), names_pattern = "(.*)_.")
```
Assign case and participant ID
```{r}
# Possible since there is no overlap in data for a given participant and case
case_df <- all_df %>%
  # Merge each row of data with each of the rows in the time map
  cross_join(time_map) %>%
  # If the given timestamp is within the time range, keep it
  # (With the associated participant ID and case)
  mutate(timestamp = if_else(between(timestamp, dt_start, dt_end),
                                  timestamp, NA), .keep = "unused") %>%
  # Drop all the timestamps which don't have a match
  drop_na()
```

## Compute PM1 Number Count
Additionally clean data a bit for future
```{r}
clean_df <- case_df %>% 
  # Calculate PM1 Count
  mutate(pm1num = bin0 + bin1 + bin2, 
  # Round date to minute and drop unnecessary columns
  date = round_date(timestamp, "minute"), .keep = "unused") %>%
  # Remove repeat readings
  distinct(case, participant_id, date, environment, .keep_all = TRUE)
```

## Compute indoor-outdoor ratios
```{r}
# Prepare dataframe to calculate ratios
df_wide <- clean_df %>% 
  # Stretch dataframe to separate indoor and outdoor readings
  pivot_wider(names_from = environment, values_from = c(pm1, pm25, pm10, pm1num,
                                                        no, co, no2)) %>%
  drop_na()
```

Compute ratios
```{r}
# Calculate indoor-outdoor ratio for each measure
df_ratio <- df_wide %>% 
  mutate(environment = "ratio",
         pm1 = pm1_indoor/pm1_outdoor, 
         pm25 = pm25_indoor/pm25_outdoor, 
         pm10 = pm10_indoor/pm10_outdoor, 
         pm1num = pm1num_indoor/pm1num_outdoor,
         co = co_indoor/co_outdoor,
         no = no_indoor/no_outdoor,
         no2 = no2_indoor/no2_outdoor,
         .keep = "unused")

# Add ratios to dataframe
modulair_df <- clean_df %>%
  # Add ratio
  rbind(df_ratio) %>%
  # Drop NAs to have clean data
  drop_na()
```

# Save Data
```{r}
save(modulair_df, file = "data_cleaned/cleaned_Olin_modulair.RData")
```

