# Aethalometer Data Cleaning
This file is to clean the Aethalometer Data for Olin deployments

## Load Libraries
```{r}
library(tidyverse)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read data
```{r}
path_to_data <- "data_raw/Olin/aethalometer/"
```

```{r}
# List all raw Aethalometer data files in Olin data
aeth_files <- list.files(path = path_to_data, recursive = TRUE)
```

```{r}
# Read all Aethalometer files
all_df <- data.frame()

for (aeth_file in aeth_files) {
  # File path
  file_path = paste0(path_to_data, aeth_file)
  
  # Get file metadata
  file_info <- file.info(file_path)
  
  # Skip empty files
  if(file_info$size > 500) {
    aeth_file_parts <- str_split_1(aeth_file, "/")
    
    
    # Get environment from file path
    env = aeth_file_parts[1]

    # Read file into dataframe
    df <- read_csv(file_path, show_col_types = FALSE) %>%
      # Select relevant variables and rename
      select(`Date / time local`, `UV BC1`, `Blue BC1`, `Green BC1`, 
             `Red BC1`, `IR BC1`, `Readable status`) %>%
      rename_with(~ tolower(gsub(" ", "_", .x)), contains("BC1")) %>%
      rename(date = `Date / time local`, status = `Readable status`) %>%
      mutate(environment = env, date = force_tz(date, tzone = "America/New_York"))

    # Combine with overall dataframe
    all_df <- rbind(all_df, df)
    
    print(paste("Read", aeth_file))
  }
}
```


## Assign case
The following section reformats the field notes to filter out the timestamps and label purifier case by participant accordingly.
```{r}
# Define lookup table to map "sham" and "hepa" to "off" and "on" respectively
case_lookup <- c(Sham = "off", Hepa = "on")
# Define table to invert the case for the second period
case_inverse <- c(off = "on", on = "off")

# Read field notes, skipping top two rows
time_map <- read_excel("data_raw/Olin/Olin_field_notes.xlsx", range = "A3:J25") %>%
  # Rename and select relevant columns
  rename(participant_id = 1, sensor_type = 10, case_1 = "Filter",
         dt_start_1 = "Air Mon 1 Visit 1", dt_end_1 = "Air Mon 1 Visit 2",
         dt_start_2 = "Air Mon 2 Visit 3", dt_end_2 = "Air Mon 2 Visit 4") %>%
  select(participant_id, case_1, starts_with("dt_"), sensor_type) %>%
  # Convert timezone to Eastern Time
  mutate(across(c(dt_start_1, dt_end_1, dt_start_2, dt_end_2), 
                ~force_tz(.x, tzone = "America/New_York")), 
         # Participant ID to numeric
         participant_id = as.numeric(participant_id),
         # Standardize case to on, off, invert the case in the second period
         case_1 = case_lookup[case_1], case_2 = case_inverse[case_1]) %>%
  # Filter for Aethalometer (misspelled in field notes)
  filter(grepl("ethylometer", sensor_type)) %>%
  select(-sensor_type) %>%
  # Stretch to create case dataframe
  pivot_longer(cols = -participant_id, names_to = c(".value"), names_pattern = "(.*)_.")
```

```{r}
clean_df <- all_df %>%
  # Filter for participants and case within time ranges (since no overlap)
  cross_join(time_map) %>% 
  mutate(time_check = between(date, dt_start, dt_end)) %>%
  filter(time_check) %>%
  # Filter out startup statuses
  filter(!grepl("Start up", status)) %>%
  # Round date to minute and drop unnecessary columns
  mutate(date = round_date(date, "minute")) %>%
  select(-c(dt_start, dt_end, status, time_check)) %>%
  # Convert all zeros to NAs
  mutate(across(contains("bc1"), ~ifelse(.x == 0, NA, .x)))
```

## Compute indoor-outdoor ratios
```{r}
# Prepare dataframe to calculate ratios
df_wide <- clean_df %>%
  # Stretch dataframe to separate indoor and outdoor readings
  pivot_wider(names_from = environment, 
              values_from = c(uv_bc1, blue_bc1, green_bc1, red_bc1, ir_bc1))
```

Compute ratios
```{r}
# Calculate indoor-outdoor ratio for each measure
df_ratio <- df_wide %>%
  mutate(environment = "ratio",
         uv_bc1 = uv_bc1_indoor/uv_bc1_outdoor, 
         blue_bc1 = blue_bc1_indoor/blue_bc1_outdoor, 
         green_bc1 = green_bc1_indoor/green_bc1_outdoor,
         red_bc1 = red_bc1_indoor/red_bc1_outdoor,
         ir_bc1 = ir_bc1_indoor/ir_bc1_outdoor,
         .keep = "unused")

# Add ratios to dataframe
aethalometer_df <- clean_df %>%
  # Add ratio
  rbind(df_ratio)
```

# Save Data
```{r}
save(aethalometer_df, file = "data_cleaned/cleaned_Olin_aethalometer.RData")
```