# Modulair-PM Data Cleaning
This file is to clean the Modulair-PM Data for Olin deployments

## Load Libraries
```{r}
library(tidyverse)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read Data
```{r}
path_to_data <- "data_raw/Olin/modpm/"

# All columns
col_list <- c("timestamp", "pm1", "pm25", "pm10", "bin0", "bin1", "bin2")
```

```{r}
# List all Mod-PM data files in Olin data
modpm_files <- list.files(path = path_to_data, recursive = TRUE)
```

## Read data
```{r}
# Initialize dataframe for all data
all_df <- data.frame()

# Loop through each deployment, read data
for (file_name in modpm_files) {
  # Split filename into its relevant parts
  file_parts <- str_split(file_name, "_")[[1]]
  
  # Read CSV, adding columns for participant ID, case, environment from filename
  df <- read_csv(paste0(path_to_data, file_name), show_col_types = FALSE) %>% 
    select(all_of(col_list)) %>%
    mutate(participant_id = as.numeric(file_parts[3]),
           case = file_parts[4],
           environment = str_remove(file_parts[5], ".csv"),
           timestamp = as.POSIXct(timestamp, tz = "America/New_York"))
  
  # Bind df to overall df
  all_df <- rbind(all_df, df)
  
  # Print that file is read
  print(paste("Completed reading", file_name))
}
```

## Time filtering
The data is manually downloaded from Quant-AQ and is filtered to the nearest day. Filter to exact time of sensor installation recorded in field notes.
```{r}
# Read field notes, skipping top two rows
time_map <- read_excel("data_raw/Olin/Olin_field_notes.xlsx", 
                       range = "A3:J25") %>%
  # Rename and select relevant columns
  rename(participant_id = 1, sensor_type = 10,
         dt_1_start = "Air Mon 1 Visit 1", dt_1_end = "Air Mon 1 Visit 2",
         dt_2_start = "Air Mon 2 Visit 3", dt_2_end = "Air Mon 2 Visit 4") %>%
  select(participant_id, sensor_type, starts_with("dt_")) %>%
  # Convert timezone to Eastern Time
  mutate(across(c(dt_1_start, dt_1_end, dt_2_start, dt_2_end), 
                ~force_tz(.x, tzone = "America/New_York"))) %>%
  # Convert participant ID to numeric, dropping invalid participants
  mutate(participant_id = as.numeric(participant_id)) %>%
  drop_na()
```

```{r}
# Create list of participants in the field notes
participant_list <- time_map %>% 
  filter(grepl("MOD-PM", sensor_type)) %>% 
  select(participant_id) %>% unique()

case_df <- data.frame()
# Loop across list of Mod-PM participants with data downloaded
for (participant in participant_list[[1]]) {
  # Filter for current participant
  current_map <- time_map %>% filter(participant_id == participant)
  
  # Filter time periods in first monitoring stage
  first_df <- all_df %>%
    filter(participant_id == participant) %>%
    filter(between(timestamp, current_map$dt_1_start, current_map$dt_1_end))
  
  # Filter time periods in second monitoring stage
  second_df <- all_df %>%
    filter(participant_id == participant) %>%
    filter(between(timestamp, current_map$dt_2_start, current_map$dt_2_end))
  
  case_df <- rbind(case_df, first_df, second_df)
}

```
## Compute PM1 Number Count
Additionally clean data a bit for future
```{r}
# Define lookup table to map "sham" and "hepa" to "off" and "on" respectively
case_lookup <- c(sham = "off", hepa = "on")

clean_df <- case_df %>% 
  # Calculate PM1 Count
  mutate(pm1num = bin0 + bin1 + bin2, 
  # Substitute case to standardize
  case = case_lookup[case],
  # Round date to minute and drop unnecessary columns
  date = round_date(timestamp, "minute"), .keep = "unused") %>%
  # Remove repeat readings
  distinct(case, participant_id, date, environment, .keep_all = TRUE)
```

## Compute indoor-outdoor ratios
```{r}
# Prepare dataframe to calculate ratios
df_wide <- clean_df %>%
  # Stretch dataframe to separate indoor and outdoor readings
  pivot_wider(names_from = environment, 
              values_from = c(pm1, pm25, pm10, pm1num)) %>%
  drop_na()
```

Compute ratios
```{r}
# Calculate indoor-outdoor ratio for each measure
df_ratio <- df_wide %>%
  mutate(environment = "ratio",
         pm1 = pm1_indoor/pm1_outdoor, pm25 = pm25_indoor/pm25_outdoor, 
         pm10 = pm10_indoor/pm10_outdoor, 
         pm1num = pm1num_indoor/pm1num_outdoor,
         .keep = "unused")

# Add ratios to dataframe
modpm_df <- clean_df %>%
  # Add ratio
  rbind(df_ratio) %>%
  # Drop NAs to have clean data
  drop_na()
```

# Save Data
```{r}
save(modpm_df, file = "data_cleaned/cleaned_Olin_modpm.RData")
```

