# CPC Data Cleaning
This file is to clean the CPC Data for Olin deployments

## Load Libraries
```{r}
library(tidyverse)
library(data.table)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read data
```{r}
path_to_data <- "data_raw/Olin/cpc/"
```

```{r}
# List all raw CPC data files in Olin data, filtering out bad and irrelevant data
cpc_files <- list.files(path = path_to_data, recursive = TRUE)
cpc_files <- grep("Pre data|bad", cpc_files, value = TRUE, invert = TRUE)
```

```{r}
# Read all CPC files - this takes a while (~ 10 min on my PC)
all_df <- data.frame()

for (cpc_file in cpc_files) {
  # File path
  file_path = paste0(path_to_data, cpc_file)
  
  # Get file metadata
  file_info <- file.info(file_path)
  
  # Skip empty files
  if(file_info$size > 500) {
    cpc_file_parts <- str_split_1(cpc_file, "/")
    
    
    # Get environment, participant ID from file path
    env = cpc_file_parts[1]
    participant = as.numeric(cpc_file_parts[2])

    # Read file into dataframe
    df <- fread(file_path, header = TRUE, skip = 13,
                select = c(1, 2, 4, 11, 12, 21, 22)) %>%
      # Merge the date and time column
      unite(col = "date", `#YY/MM/DD`, `HR:MN:SC`, sep = " ") %>%
      # Convert to datetime object and insert environment, participant ID
      mutate(date = as.POSIXct(date, format = "%y/%m/%d %H:%M:%S", 
                               tz = "America/New_York"),
             environment = env, participant_id = participant)

    # Combine with overall dataframe
    all_df <- rbind(all_df, df)
    
    print(paste("Read", cpc_file))
  }
}
```

## Assign case
The following section reformats the field notes to filter out the timestamps and label purifier case by participant accordingly.
```{r}
# CPC participants
cpc_participants <- unique(all_df$participant_id)

# Define lookup table to map "sham" and "hepa" to "off" and "on" respectively
case_lookup <- c(Sham = "off", Hepa = "on")
# Define table to invert the case for the second period
case_inverse <- c(off = "on", on = "off")


# Read field notes, skipping top two rows
time_map <- read_excel("data_raw/Olin/Olin_field_notes.xlsx", range = "A3:I25") %>%
  # Rename and select relevant columns
  rename(participant_id = 1, case_1 = "Filter",
         dt_start_1 = "Air Mon 1 Visit 1", dt_end_1 = "Air Mon 1 Visit 2",
         dt_start_2 = "Air Mon 2 Visit 3", dt_end_2 = "Air Mon 2 Visit 4") %>%
  select(participant_id, case_1, starts_with("dt_")) %>%
  # Convert timezone to Eastern Time
  mutate(across(c(dt_start_1, dt_end_1, dt_start_2, dt_end_2), 
                ~force_tz(.x, tzone = "America/New_York")), 
         # Participant ID to numeric
         participant_id = as.numeric(participant_id),
         # Standardize case to on, off, invert the case in the second period
         case_1 = case_lookup[case_1], case_2 = case_inverse[case_1]) %>%
  # Filter for Modulair participants
  filter(participant_id %in% cpc_participants) %>%
  # Stretch to create case dataframe
  pivot_longer(cols = -participant_id, names_to = c(".value"), names_pattern = "(.*)_.")
```

### Helper function
```{r}
# Filter timestamps and include whether hepa purifier is on or off
include_case <- function(df, participant) {
  # Filter by current participant
  cd <- time_map %>% filter(participant_id == participant)
  
  df %>%
    # Select current participant
    filter(participant_id == participant) %>%
    # Add case column
    mutate(case = case_when(
      # For off periods
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start[cd$case == "off"], 
              cd$dt_end[cd$case == "off"]) ~ 'off',
      # For on periods
      between(date, cd$dt_start[cd$case == "on"], 
              cd$dt_end[cd$case == "on"]) ~ 'on'))
}
```

### Assign case for each participant
```{r}
case_df <- data.frame()

for (participant in cpc_participants) {
  # Filter by date for off and on (see helper function)
  df <- include_case(all_df, participant)
  # Append to overall dataframe, removing readings from invalid times
  case_df <- rbind(case_df, df) %>% drop_na()
  print(paste("Assigned case to", participant))
}
```

## Clean data
```{r}
clean_df <- case_df %>%
  # Pump flow needs to be greater than 250 for proper functioning
  filter(smpflow > 250) %>%
  # Drop pump flow variable
  select(-smpflow) %>%
  # Fill count has to be zero for correct data collection
  filter(fillcnt == 0) %>%
  # Drop pump flow variable
  select(-fillcnt) %>%
  # Error number has to be zero for correct data collection
  filter(err_num == 0) %>%
  # Drop pump flow variable
  select(-err_num) %>%
  # Convert concent to float
  mutate(concent = as.numeric(concent)) %>%
  # Filter out concent values that are < 100 (this can never be true IRL)
  filter(concent > 100) %>%
  # Remove repeat readings
  distinct(participant_id, date, environment, .keep_all = TRUE) %>%
  # Drop temperature values (since not required for now)
  select(-inlttmp) %>%
  # Remove all filtered-out values
  drop_na()
  
```

## Compute indoor-outdoor ratios
CPC ratios are calculated to nearest minute to match Modulair/Mod-PM resolution.
```{r}
# Prepare dataframe to calculate ratios - takes about a minute to run
df_round <- clean_df %>%
  # Round values to nearest minute and calculate mean
  mutate(date_round = round_date(date, "1 minute")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, date_round) %>%
  # Calculate mean per minute
  summarize(concent = mean(concent), .groups = "drop")

```

```{r}
# Stretch dataframe to separate indoor and outdoor readings
df_wide <- df_round %>%
  pivot_wider(names_from = environment, values_from = concent) %>%
  drop_na()
```


Compute ratios
```{r}
# Calculate indoor-outdoor ratio for each measure
df_ratio <- df_wide %>%
  mutate(environment = "ratio", concent = indoor/outdoor, .keep = "unused") %>%
  rename(date = date_round)

# Add ratios to dataframe
cpc_df <- clean_df %>%
  # Add ratio
  rbind(df_ratio) %>%
  # Drop NAs to have clean data
  drop_na()
```



# Save data
```{r}
save(cpc_df, file = "data_cleaned/cleaned_Olin_cpc.RData")
```

