# Modulair Data Cleaning
This file is to clean the Modulair Data for SCOPE deployments

## Load Libraries
```{r}
library(tidyverse)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read data
```{r}
path_to_data <- "data_raw/SCOPE/"

# All deployments
deployment_ids <- 1:4

# All columns
col_list <- c("timestamp", "pm1", "pm25", "pm10", "no", "co", "no2")
```

```{r}
# Initialize dataframe for all data
all_df <- data.frame()

# Loop through each deployment, read data
for (deployment in deployment_ids) {
  path <- paste0(path_to_data, "deployment_", deployment, "/")
  # Read indoor data
  indoor <- read_csv(paste0(path, "MOD_indoor.csv"), show_col_types = FALSE) %>%
    select(col_list) %>% mutate(environment = "indoor")
  
  # Read outdoor data
  outdoor <- read_csv(paste0(path, "MOD_outdoor.csv"), show_col_types = FALSE) %>%
    select(col_list) %>% mutate(environment = "outdoor")
  
  # Combine indoor and outdoor data, inserting deployment ID
  joined <- rbind(indoor, outdoor) %>% mutate(dep_id = deployment)
  
  all_df <- rbind(all_df, joined)
}
```


## Assign case
Each air quality reading was either when the purifier was running, not running, or outside the study period. The following section reformats the field notes to filter out the timestamps and label purifier case by deployment accordingly.
### Get purifier off/on lookup table
```{r}
# Read field notes, skipping empty top row
time_map <- read_csv(paste0(path_to_data, "SCOPE_field_notes.csv"), skip = 1)

# Read out first leg of all deployments
run_1 <- time_map %>% 
  select("Deployment #", "Start", "Status...3", "Change") %>%
  rename("dt_start" = "Start", "case" = "Status...3", "dt_end" = "Change")

# Read out second leg of all deployments
run_2 <- time_map %>% 
  select("Deployment #", "Change", "Status...5", "End") %>%
  rename("dt_start" = "Change", "case" = "Status...5", "dt_end" = "End")

# Join them, converting dates to datetime format
rbind(run_1, run_2) %>% 
  rename("dep_id" = "Deployment #") %>%
  mutate(across(c(dt_start, dt_end), ~as.POSIXct(.x, 
                                                 format = "%m/%d/%Y %I:%M %p",
                                                 tz = "America/New_York"))) %>%
  mutate(case = gsub("HEPA ", "", case)) -> time_map
```

### Helper function
```{r}
# Filter timestamps and include whether hepa purifier is on or off
include_case <- function(df, deployment) {
  # Filter by current deployment
  cd <- time_map %>% filter(dep_id == deployment)
  
  df %>%
    # Select current deployment
    filter(dep_id == deployment) %>%
    # Add case column
    mutate(case = case_when(
      # For off periods
      # Note: 'cd' is the table 'current_dates' in function call
      between(timestamp, cd$dt_start[cd$case == "off"], 
              cd$dt_end[cd$case == "off"]) ~ 'off',
      # For on periods
      between(timestamp, cd$dt_start[cd$case == "on"], 
              cd$dt_end[cd$case == "on"]) ~ 'on'))
}
```

### Assign case for each deployment
```{r}
case_df <- data.frame()

for (deployment in deployment_ids) {
  # Filter by date for off and on (see helper function)
  df <- include_case(all_df, deployment)
  # Append to overall dataframe, removing readings from invalid times
  case_df <- rbind(case_df, df) %>% drop_na()
}
```

## Compute indoor-outdoor ratios
```{r}
# Prepare dataframe to calculate ratios
df_wide <- case_df %>% 
  # Round timestamps to nearest minute, rename column to date (for plotting)
  mutate(date = round_date(timestamp, "minute"), .keep = "unused") %>%
  # Remove repeat readings
  distinct(case, dep_id, date, environment, .keep_all = TRUE) %>%
  # Stretch dataframe to separate indoor and outdoor readings
  pivot_wider(names_from = environment, values_from = pm1:no2) 
```

Compute ratios
```{r}
df_ratio <- select(df_wide, ends_with("indoor"))/
    select(df_wide, ends_with("outdoor"))

# Rename computed columns
colnames(df_ratio) <- sub("_indoor", "_ratio", colnames(df_ratio))

# Add ratios to dataframe and reshape to original format
modulair_df <- df_wide %>%
  # Add ratio
  cbind(df_ratio) %>%
  # Reshape, delimiter is _
  pivot_longer(pm1_indoor:no2_ratio, names_to = c(".value", "environment"), 
               names_pattern = "(.*)_(.*)") %>%
  # Drop NAs to have clean data
  drop_na()
```

# Save Data
```{r}
save(main_df, file = "cleaning/cleaned_SCOPE_modulair.RData")
```

