# CPC Data Cleaning
This file is to clean the CPC Data for SCOPE deployments

## Load Libraries
```{r}
library(tidyverse)
library(data.table)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read data
```{r}
path_to_data <- "data_raw/SCOPE/"

# All deployments
deployment_ids <- 1:4
```

```{r}
# List all raw CPC data files in SCOPE data
all_files <- list.files(path = path_to_data, recursive = TRUE)
cpc_files <- grep("cpc_data", all_files, value = TRUE)
```

```{r}
# Read all CPC files
all_df <- data.frame()

for (cpc_file in cpc_files) {
  # File path
  file_path = paste0(path_to_data, cpc_file)
  
  # Get file metadata
  file_info <- file.info(file_path)
  
  # Skip empty files
  if(file_info$size > 500) {
    # Get deployment ID from file path
    deployment = as.integer(substr(cpc_file, 12, 12))
    
    # Read file into dataframe, adding deployment ID
    df <- fread(file_path, header = TRUE, skip = 13, 
                select = c(1, 2, 4, 11, 12, 20, 21)) %>% mutate(dep_id = deployment)
    
    # Combine with overall dataframe
    all_df <- rbind(all_df, df)
  }
}
```

## Add environment to data
The CPC was connected to a valve that switched between indoor and outdoor at regular intervals. This section retrieves the processed valve-switching data and merges it with the CPC data.
```{r}
# Warning: this code chunk may seem simple but takes FOREVER.
# Have patience. It deals with datetimes.
merged_date_df <- all_df %>%
  # Merge the date and time column
  unite(col = "date", `#YY/MM/DD`, `HR:MN:SC`, sep = " ") %>%
  # Convert to datetime object
  mutate(date = as.POSIXct(date, format = "%y/%m/%d %H:%M:%S", 
                           tz = "America/New_York"))
```

Load environment data from valve switching
```{r}
# This file was created by `SCOPE_valve_processing.Rmd`
env_df <- read_csv(paste0(path_to_data, "valve_data/valve_combined.csv"))
```
Merge CPC and valve switching (environment) data
```{r}
include_env_df <- merged_date_df %>%
  inner_join(env_df, by = join_by(date == datetime, dep_id))
```

## Assign case
Each air quality reading was either when the purifier was running, not running, or outside the study period. The following section reformats the field notes to filter out the timestamps and label purifier case by deployment accordingly.
### Get purifier off/on lookup table
```{r}
# Read field notes, skipping empty top row
time_map <- read_csv(paste0(path_to_data, "SCOPE_field_notes.csv"), skip = 1)

# Read out first leg of all deployments
run_1 <- time_map %>% 
  select("Deployment #", "Start", "Status...3", "Change") %>%
  rename("dt_start" = "Start", "case" = "Status...3", "dt_end" = "Change")

# Read out second leg of all deployments
run_2 <- time_map %>% 
  select("Deployment #", "Change", "Status...5", "End") %>%
  rename("dt_start" = "Change", "case" = "Status...5", "dt_end" = "End")

# Join them, converting dates to datetime format
rbind(run_1, run_2) %>% 
  rename("dep_id" = "Deployment #") %>%
  mutate(across(c(dt_start, dt_end), ~as.POSIXct(.x, 
                                                 format = "%m/%d/%Y %I:%M %p",
                                                 tz = "America/New_York"))) %>%
  mutate(case = gsub("HEPA ", "", case)) -> time_map
```

### Helper function
```{r}
# Filter timestamps and include whether hepa purifier is on or off
include_case <- function(df, deployment) {
  # Filter by current deployment
  cd <- time_map %>% filter(dep_id == deployment)
  
  df %>%
    # Select current deployment
    filter(dep_id == deployment) %>%
    # Add case column
    mutate(case = case_when(
      # For off periods
      # Note: 'cd' is the table 'current_dates' in function call
      between(date, cd$dt_start[cd$case == "off"], 
              cd$dt_end[cd$case == "off"]) ~ 'off',
      # For on periods
      between(date, cd$dt_start[cd$case == "on"], 
              cd$dt_end[cd$case == "on"]) ~ 'on'))
}
```

### Assign case for each deployment
```{r}
case_df <- data.frame()

for (deployment in deployment_ids) {
  # Filter by date for off and on (see helper function)
  df <- include_case(include_env_df, deployment)
  # Append to overall dataframe, removing readings from invalid times
  case_df <- rbind(case_df, df) %>% drop_na()
}
```

## Clean data
```{r}
clean_df <- case_df %>%
  # Pump flow needs to be greater than 250 for proper functioning
  filter(smpflow > 250) %>%
  # Drop pump flow variable
  select(-smpflow) %>%
  # Fill count has to be zero for correct data collection
  filter(fillcnt == 0) %>%
  # Drop pump flow variable
  select(-fillcnt) %>%
  # Error numeber has to be zero for correct data collection
  filter(err_num == 0) %>%
  # Drop pump flow variable
  select(-err_num) %>%
  # Filter out concent values that are < 100 (this can never be true IRL)
  filter(concent > 100) %>%
  # Remove repeat readings
  distinct(dep_id, date, environment, .keep_all = TRUE) %>%
  # Drop temperature values (since not required for now)
  select(-inlttmp) %>%
  # Remove all filtered-out values
  drop_na()
  
```

## Calculate indoor-outdoor ratios
Since there is no concurrent indoor and outdoor data, ratios are calculated over the average of indoor and outdoor values in one valve switch cycle. The cycle starts with outdoor to help account for infiltration effects.
```{r}
ratio_df <- data.frame()

# Calculate by deployment ID
for (deployment in deployment_ids) {
  # Create section variable that increments every time the environment switches
  with_section <- clean_df %>%
    filter(dep_id == deployment) %>%
    mutate(section = cumsum(environment != lag(environment, default = "")))
  
  # Ensure cycle starts with outdoor - if start indoor, filter out the top section
  if (with_section$environment[1] == "indoor") {
    with_section <- with_section %>% filter(section > 1)
    # If start outdoor, increment by 1 to match previous if case (for floor division)
  }  else {
    with_section <- with_section %>% mutate(section = section + 1)
  }
  
  # Make indoor-outdoor pair by floor dividing 
  with_section <- with_section %>% mutate(section = section %/% 2)
  
  # Get the starting timestamp and case of each section, for later
  start_timestamp <- with_section %>%
    group_by(section) %>%
    slice(1) %>%
    ungroup() %>%
    select(section, date, case)
  
  # Calculate mean indoor and outdoor per section, and then calculate ratio 
  ratio_by_dep <- with_section %>%
    # Calculate mean for each section
    group_by(section, environment) %>%
    summarize(concent = mean(concent), .groups = "drop") %>%
    # Calculate indoor-outdoor ratio
    pivot_wider(names_from = environment, values_from = concent) %>%
    mutate(ratio = indoor/outdoor, .keep = "unused") %>%
    # Rename to variable name to match with original
    rename(concent = ratio) %>%
    # Add in timestamps to sync with section, and drop section
    inner_join(start_timestamp, by = join_by(section)) %>% 
    select(-section) %>%
    mutate(dep_id = deployment)
    
  ratio_df <- rbind(ratio_df, ratio_by_dep)
}

ratio_df <- ratio_df %>% mutate(environment = "ratio") %>% drop_na()
```

```{r}
# Append the ratio data to the raw data
cpc_df <- rbind(clean_df, ratio_df)
```

# Save data
```{r}
save(cpc_df, file = "data_cleaned/cleaned_SCOPE_cpc.RData")
```

