# CPC Data Cleaning
This file cleans the CPC Data for Tufts deployments

## Load Libraries
```{r}
library(tidyverse)
library(data.table)
library(readxl)
```

Set working directory
```{r, setup, include=FALSE}
working_dir <- "HAFTRAP-summary"
knitr::opts_knit$set(root.dir = paste0("~/", working_dir))
```

Check for working directory
```{r}
# Check file path
if (!endsWith(getwd(), working_dir)) {
  stop("Incorrect working directory")
}
```

## Read data
```{r}
path_to_data <- "data_raw/Tufts/cpc/"
```

```{r}
# List all raw CPC data files in Tufts data, filtering out irrelevant files
cpc_files <- list.files(path = path_to_data, recursive = TRUE)
cpc_files <- grep("xlsx|txt", cpc_files, value = TRUE, invert = TRUE)
```

```{r}
# Read all CPC files - this takes a while (~ 5 min on my PC)
all_df <- data.frame()

for (cpc_file in cpc_files) {
  # File path
  file_path = paste0(path_to_data, cpc_file)
  
  # Get file metadata
  file_info <- file.info(file_path)
  
  # Skip empty files
  if(file_info$size > 500) {
    cpc_file_parts <- str_split_1(cpc_file, "/")
    
    # Get participant ID
    participant_parts <- str_split_1(cpc_file_parts[1], " ")
    participant <- as.integer(participant_parts[3])
    
    # Get environment  
    env_period_str <- str_to_lower(cpc_file_parts[2])
    
    env <- NA
    if (grepl("in", env_period_str)) {
      env <- "indoor"
    } else if (grepl("out", env_period_str)){
      env <- "outdoor"
    }
    
    # Get period, default is 1
    per <- 1
    if (grepl("2", env_period_str)) {
      per <- 2
    }
    
    # Read file into dataframe
    df <- fread(file_path, header = TRUE, skip = 5) %>%
      # Merge the date and time column
      unite(col = "date", "Date", "Time", sep = " ") %>%
      # Convert to datetime object and insert environment, participant ID
      mutate(date = as.POSIXct(date, format = "%Y/%m/%d %H:%M:%S", 
                               tz = "America/New_York"),
             environment = env, participant_id = participant, period = per)

    # Combine with overall dataframe
    all_df <- rbind(all_df, df)
    
    print(paste("Read", cpc_file))
  }
}
```

```{r}
# Clean data by removing duplicates, filtering out failed or invalid scans
clean_df <- all_df %>%
  # Remove repeat readings
  distinct(participant_id, date, environment, period, .keep_all = TRUE) %>%
  # Rename relevant columns to be more descriptive
  rename(concent = Concentration, cpc_count = Count, status_flag = `Status Flags`) %>%
  # Status flag of 0 is successful scan - filter out all others, as well as invalid data
  filter(status_flag == "0", concent > 100)
```


Table mapping periods to sham or HEPA case
```{r}
# Read summary table of times
time_map <- read_excel("data_raw/Tufts/Tufts_field_notes_formatted.xlsx")
```

Match sham and HEPA case to periods
```{r}
case_df <- inner_join(clean_df, time_map, by = join_by(participant_id, period)) %>%
  # Convert sham and hepa to off and on respectively, for standardization
  mutate(case = case_when(case == "sham" ~ "off", case == "hepa" ~ "on")) %>%
  # Some post-processing - select relevant columns, drop invalid values
  select(date, participant_id, case, environment, concent) %>%
  drop_na()
```

## Compute indoor-outdoor ratios
CPC ratios are calculated to nearest minute to match Modulair/Mod-PM resolution.
```{r}
# Prepare dataframe to calculate ratios - takes about a minute to run
df_round <- case_df %>%
  # Round values to nearest minute and calculate mean
  mutate(date_round = round_date(date, "1 minute")) %>%
  # Group by the rounded datetime
  group_by(case, environment, participant_id, date_round) %>%
  # Calculate mean per minute
  summarize(concent = mean(concent), .groups = "drop")
```

```{r}
# Stretch dataframe to separate indoor and outdoor readings
df_wide <- df_round %>%
  pivot_wider(names_from = environment, values_from = concent) %>%
  drop_na()
```


Compute ratios
```{r}
# Calculate indoor-outdoor ratio for each measure
df_ratio <- df_wide %>%
  mutate(environment = "ratio", concent = indoor/outdoor, .keep = "unused") %>%
  rename(date = date_round)

# Add ratios to dataframe
cpc_df <- case_df %>%
  # Add ratio
  rbind(df_ratio) %>%
  # Drop NAs to have clean data
  drop_na()
```

# Save data
```{r}
save(cpc_df, file = "data_cleaned/cleaned_Tufts_cpc.RData")
```


---- End of Cleaning Script ----

## Plotting before/after cleaning
```{r}
# Round raw data to 1 minute for quicker plotting
plot_df <- all_df %>%
  # Round values to nearest minute and calculate mean
  mutate(date_round = round_date(date, "1 minute")) %>%
  # Group by the rounded datetime
  group_by(environment, participant_id, date_round, `Status Flags`) %>%
  # Calculate mean per minute
  summarize(Concentration = mean(Concentration), .groups = "drop") %>%
  mutate(participant_id = as.factor(participant_id))
```

```{r}
plot_df %>%
ggplot(aes(x = date_round, y = Concentration, color = participant_id)) +
  geom_point(size = 0.1) +
  facet_wrap(~environment, nrow = 2)
```

```{r}
# Plot cleaned data using df_round from above 
# (df_round is post-filtered data rounded to 1 minute)
df_round %>%
  mutate(participant_id = as.factor(participant_id)) %>%
  ggplot(aes(x = date_round, y = concent, color = participant_id)) +
    geom_point(size = 0.1) +
    facet_wrap(~environment, nrow = 2)
```

